---
title: "**Missing Data Report - Water Quality - Nitrate**"
output:
  html_document:
    toc: true
    code_folding: hide
    highlight: tango
    self_contained: true
---


This is a report for water quality with data stored in the WIMS table.

?Set the parameters, and then run the report. The data will be pulled from the table on Databricks, and a report will be generated in the `docs` folder. ?Additionally, a `csv` file containing the missing data by site will be generated in the `outputs/tables` folder.

`st` = network, set to "LS"

`dc` = det code (should be det code but based on the LSN data available it is det_desc)

`fn` = prefix for output file names

`fnn` = the name of the variable, used in file names and figures.

`years` = which years to get missing data for

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  attr.source = ".numberLines")



#load required libraries
library(brickster)

```

```{r}
##to load the data from databricks
warehouses <- db_sql_warehouse_list()

con <- dbConnect(
  drv = DatabricksSQL(),
  warehouse_id = "0a86ae40313bb7db",
  catalog = "prd_dash_lab")


LSN_data <- dbGetQuery(con,
                       "SELECT * FROM prd_dash_lab.seda_unrestricted.ard_lsn_tbl_sample_wims")
```



```{r}
### Set and save the parameters to use for the report
### Select one, comment out the rest
## LSN:
param <- list(st="LS", dc="Ammoniacal Nitrogen, Filtered as N", fn="lsn_wims_", fnn="Nitrogen", years=2024:2025)
#param <- list(st="LS", dc="Orthophosphate, Filtered as P", fn="lsn_wims_", fnn="Phosphate", years=2024:2025)
#param <- list(st="LS", dc="Carbon, Organic, Dissolved as C :- {DOC}", fn="lsn_wims_", fnn="DoC", years=2024:2025)	
#param <- list(st="LS", dc="Temperature of Water", fn="lsn_wims_", fnn="Temperature", years=2024:2025)
```

```{r}
# Load libraries
library(data.table) # data management
library(DBI)
library(sf)
library(dplyr)
library(DT)
library(lubridate)
library(tidyr)
# Run the code to process missing data

source("../LSN_missingness_WIMS.R")
source("../missing_data_plot_functions.r")
source("../missing_data_functions.r")
```

## **Missing data per unique site**

Analysis of missing data for det desc: Ammoniacal Nitrogen, Filtered as N, for the LSN network, which is sampled once per month.

Missing data is determined using the variable: meas_result

The table shows the number of sites sampled per year, although not all were sampled. “Planned visits” are the number of sites that were intended to be sampled that year. “Sampled at least once” is the number of sites visited at least once. “Missing 1 month or more” is the number of sites where there is some missing data for that year. “Planned, never sampled” is the number of sites that were not sampled at all that year.

```{r}
### Missing sites table
# Planned sites with 1 missing month or more
LSN_missing_1more <- lapply(1:length(param$years), function(x) nrow(LSN_wide[[x]][missing_count > 0])) |> unlist()
# Combine into 1 table and rearrange column order
tbl_data_missing <- cbind(LSN_sites_table, LSN_missing_1more)[, c(1:4, 6, 5)]
snames <- c("Year", "Original design", "Planned visits", "Sampled at least once", "Missing 1 month or more", "Planned, never sampled")
colnames(tbl_data_missing) <- snames
# Convert to DT for
dt_site_table <- datatable(tbl_data_missing, options=list(dom='t', paging=FALSE), rownames=FALSE)

dt_site_table
```



“Missingness” is calculated against the number of planned visit sites in any given year. This includes sites that were retired or were/are replacement sites.

Histogram showing percentage of missing data by site:

```{r}
# Histogram
par(mar=c(4.5,4.5,1,1))
hist(LSN_site_miss_table[, missing_count_pc], breaks=seq(0, 100, 10), col=colorRampPalette(c("#405fc7", "#a5b9fa"))(11), xlab="Missing Data %", main="", axes=FALSE)
axis(1, at=seq(0, 100, 10), las=1)
axis(2, at=seq(0, 1600, 100), las=1)
#text(x=0, y=-10, "None", xpd=TRUE, srt=90)

```

# **National level**

The barplots show the level of missing data at a national level per month and per year (as percentage):


```{r}
# Round up value to even
lay_n <- round_up_even(length(param$years))
# Plot
par(mar=c(4,4,4,4))
layout(matrix(1:lay_n, ncol=2, byrow=TRUE))
# Yearly change by month
for(i in 1:length(param$years)) {
    plot_miss_nat_yearly_month(x=LSN_missing_nat_percent[[i]] |> unlist(), title=paste0("Monthly (", param$years[i], ")"), names.arg=format(ISOdate(2010,1:12, 1),"%b"))
}
```

And the mean monthly values:

```{r}
# Average monthly
par(mar=c(4,4,4,4))
plot_miss_nat_yearly_month(x=LSN_missing_nat_totals_pc, title=paste0("Mean Monthly (", yr[1], " - ", yr[2], ")"), col=colorRampPalette(c("skyblue2", "skyblue4"))(12))

```

The mean level (%) of missing data per year is:

```{r}
# Format table to show mean misisng data per year
tbl_data_year_missing <- data.frame(Year = param$years, Missing = unlist(lapply(LSN_missing_nat_percent, \(x) round(mean(unlist(x)), 1))))
dt_year_missing <- datatable(tbl_data_year_missing, options = list(dom = 't', paging = FALSE, autoWidth = TRUE), rownames = FALSE, width = "20")

dt_year_missing
```

The mean level (%) of missing data per month across 2023 - 2025 is:

```{r}
# Format table to show mean misisng data per year
tbl_data_month_missing <- data.frame(Month = format(ISOdate(2010,1:12, 1), "%B"), Missing = round(LSN_missing_nat_totals_pc, 1))
dt_month_missing <- DT::datatable(tbl_data_month_missing, options = list(dom = 't', paging = FALSE, autoWidth = TRUE), rownames = FALSE, width = "20")

dt_month_missing
```

## **Littles MCAR Test**

National level missing data was tested to see whether it should be considered Missing Completely at Random (MCAR). This test helps determine whether the pattern of missing measurements appears systematic or essentially unrelated to the data. These results should be considered as supporting evidence of the type of missing data (if no results are shown, it means test was not run due to too much missing data):

```{r}
# Skip this for variables with too much missing data (it fails)
skip_mcar <- FALSE
if(param$st=="LS" & param$fnn=="Nitrogen") {skip_mcar <- TRUE} 
if(skip_mcar==FALSE) {   
# Run test for each year
mcar_res <- lapply(LSN_wide, \(x) littles_mcar_test(x[, -c(1, 2)]))
# Convert results
res_dt <- rbindlist(lapply(mcar_res, as.data.table))
res_dt[, `:=`(statistic = as.integer(round(statistic)), p.value = round(p.value, 4))]
# As DT for report display
res_dt[, year := param$years]
setcolorder(res_dt, "year")
# Change to DT for display
res_dt <- DT::datatable(res_dt, options = list(dom = 't', paging = FALSE, autoWidth = FALSE), rownames = FALSE, width = "auto")}

```

For datasets with very high levels of missing data the test may not be estimable and therefore no results is presented. P values that are not significant (greater than 0.05) suggest that the data could be considered MCAR. Where the p value is significant (less than 0.05), the data can not be considered MCAR. The data will either be Missing at Random (MAR), or Missing Not at Random (MNAR).


# **Regional level**

The heatmaps show the level of missing data at a regional level (e.g. by river basin district) per month and per year (as percentage):

```{r}
# Plot
par(mar=c(4,4,4,4))
layout(matrix(1:lay_n, ncol=2, byrow=TRUE))
# Heatmaps
for(i in 1:length(param$years)) {
    plot_miss_heatmap(x=LSN_missing_rbd_percent[[i]][, .SD, .SDcols=LSN_wide_names[[i]]] |> as.matrix(), xlabs=format(ISOdate(2010,1:12, 1),"%b"), ylabs=LSN_missing_rbd_percent[[i]][, ea_rbd], title=paste0("Monthly (", param$years[i], ")"), lta=0.5)
}
```

The mean level (%) of missing data per RBD per year is:

```{r}
# Format table/results
LSN_missing_rbd_yearly_mean <- lapply(LSN_missing_rbd_percent, \(x) round(x[, yearly_mean], 1)) |> do.call(cbind, args=_)
colnames(LSN_missing_rbd_yearly_mean) <- param$year
LSN_missing_rbd_yearly_mean <- data.table(RBD = as.character(LSN_missing_rbd_percent[[1]][, ea_rbd]), LSN_missing_rbd_yearly_mean)
# DT for pretty table
LSN_missing_rbd_yearly_mean_dt <- DT::datatable(LSN_missing_rbd_yearly_mean, options = list(dom = 't', paging = FALSE, autoWidth = TRUE), rownames = FALSE, width = "20")

LSN_missing_rbd_yearly_mean_dt
```

# **Site level**

The maps show the level of missing data at a site level, with the number of missing months per site per year:

```{r, results='hide'}
### Create SF objects to plot maps
# Add geometry
lapply(1:length(param$years), \(x) LSN_wide[[x]][LSN_sites_filtered, geometry := i.geometry, on="network_id"])
# add percent missing data
lapply(1:length(param$years), \(x) LSN_wide[[x]][, missing_count_pc := missing_count / 12 * 100])
# Create sf object for plotting (total missing per year)
LSN_wide_sf <- lapply(1:length(param$years), \(x) st_as_sf(LSN_wide[[x]][, .(missing_count, missing_count_pc, geometry)]))
# create categories
months <- format(ISOdate(2010,1:12,1), "%b")
missing_cat <- lapply(1:length(param$years), \(x) apply(LSN_wide[[x]][, .SD, .SDcols=LSN_wide_names[[x]]], 2, \(x) ifelse(is.na(x), "Missing", "Not Missing")))
# rename
for(i in 1:length(param$years)) { colnames(missing_cat[[i]]) <- months}
# Add to sf
LSN_wide_sf <- lapply(1:length(param$years), \(x) cbind(LSN_wide_sf[[x]], missing_cat[[x]]))

```


```{r}
### Plot
# Create colour maps (custom function)
mcols <- vector("list", length(LSN_wide_sf))
for(i in 1:length(LSN_wide_sf)) {mcols[[i]] <- col_map(x=LSN_wide_sf[[i]]$missing_count)}
lcol <- hcl.colors(13, "YlOrRd", rev = TRUE) # needs to match what went into the colour map
#
par(mar=c(2, 4, 3, 3))
layout(matrix(1:lay_n, ncol=2, byrow=TRUE))
brks <- 0:12
# Maps
for(i in 1:length(param$years)) {
    plot(st_geometry(sh_eng), lwd = 2, axes=TRUE, main=param$years[i])
    plot(LSN_wide_sf[[i]]["missing_count"], pch=21, col="black", bg=mcols[[i]], add=TRUE)
    legend(x = par("usr")[2], y = mean(par("usr")[3:4]), legend=brks, pt.bg=lcol, pch=21, pt.cex=1.5, xpd = NA, bty = "n", yjust = 0.5)
}

```

## **Missing data by site**

The sites with the highest level of missing data are (only shows the first 20 rows):

```{r}
sites_miss_dt <- datatable(LSN_site_miss_table[1:20,], colnames=c("ID", "Site status", "RBD", "No. Years", "Missing data", "Missing data %"))

sites_miss_dt
```

A full table of results can be found in the `outputs/tables` folder


















